Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-06-03 manuscript.Rmd
Started eLife revisions

---
title: "Different rules for binocular combination of luminance flicker in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker"
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  pdf_document:
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

processdata <- 0  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data
# 2 - average data across participants, do bootstrapping and modelling
# 3 - download all raw data and analyse (requires 48GB of storage)

@@

         # and delete the tar file to save storage space
         file.remove(paste0(rawdir,'P',100+participant,'.tar'))
       }

@@

    save(file=paste0(datadir,'P',150+participant,'summary.RData'),list=c('cleanmeansE','meanspec'))

@@

# settings for sampler 
numSavedSteps <- 100000    # total number of MCMC steps
adaptSteps <- 1000  # Number of steps to "tune" the samplers
burnInSteps <- 2000 
thinSteps <- 10
nChains <- 8
ncores <- detectCores()/2  # because of multi-threading, this is usually double the true number of physical cores, so we halve the number as Stan can only use physical cores
options(mc.cores=ncores)

>>>>

---
title: "Different rules for binocular combination of luminance flicker in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker"
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document2: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

processdata <- 3  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data
# 2 - average data across participants, do bootstrapping and modelling
# 3 - download all raw data and analyse (requires 48GB of storage)

@@

         # and delete the tar file to save storage space
         file.remove(paste0(rawdir,'P',100+participant,'.tar'))
       }
      }

@@

    save(file=paste0(datadir,'P',150+participant,'summary.RData'),list=c('cleanmeansE','meanspec'))
    }

@@

# settings for sampler 
numSavedSteps <- 100000    # total number of MCMC steps
adaptSteps <- 1000  # Number of steps to "tune" the samplers
burnInSteps <- 2000 
thinSteps <- 10
nChains <- 8
ncores <- 8  # because of multi-threading, this is usually double the true number of physical cores, so we halve the number as Stan can only use physical cores
options(mc.cores=ncores)

### 2023-06-03 manuscript.tex
Created rmd file specific for Nature and added Nature referencing style.
Started eLife revisions

Figure \ref{fig:pupildata}c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A \(2\times5\) repeated measures \(ANOVA^2_{circ}\) (Baker, 2021) comparing these conditions revealed a significant main effect of target contrast (F(8,580) = 16.79, \(p\) \textless{} 0.001), a significant effect of condition (F(2,580) = 11.04, \(p\) \textless{} 0.001), and a significant interaction (F(8,580) = 56.25, \(p\) \textless{} 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48\%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(8,232) = 3.03, \(p\) \textless{} 0.003).

@@

Finally, we calculated the ratio of binocular to monocular responses across the three data types from Experiment 1. Figure \ref{fig:BSratios} shows that these ratios are approximately \(\sqrt2\) across the low-to-intermediate contrast range for all three data types. At higher contrasts, we see ratios of 2 or higher for the EEG data, but much weaker ratios near 1 for the pupillometry data. Note that the ratios here are calculated on a per-participant basis and then averaged, rather than being the ratios of the average values shown in Figures \ref{fig:pupildata} and \ref{fig:EEGdata}. A \(3 \times 5\) repeated measures ANOVA on the logarithmic (dB) ratios found a main effect of contrast (F(3.08,89.28) = 4.53, \emph{p} \textless{} 0.002), no effect of data modality (F(2,58) = 0.75, \emph{p} = 0.48), but a highly significant interaction (F(5.54,160.67) = 3.84, \emph{p} \textless{} \ensuremath{3\times 10^{-4}}).

@@

\caption{\label{tab:paramtable}Summary of median parameter values.}
\centering
\begin{tabular}[t]{l|c|c|c|c}
\hline
Data set & Z & k & w & Rmax\\
\hline
Pupillometry & 2.37 & 0.01 & 0.67 & 0.00023\\
\hline
EEG 1F & 2.59 & 0.15 & 0.02 & 0.0026\\
\hline
EEG 2F & 2.21 & 0.06 & 0.01 & 0.00404\\
\hline
Matching & 0.30 & 5.07 & 0.09 & -\\
\hline
\end{tabular}
\end{table}

@@

>>>>

Figure \ref{fig:pupildata}c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A \(2\times5\) repeated measures \(ANOVA^2_{circ}\) (Baker, 2021) comparing these conditions revealed a significant main effect of target contrast (F(8,580) = 16.79, \(p\) \textless{} 0.001), a significant effect of condition (F(2,580) = 11.04, \(p\) \textless{} 0.001), and a significant interaction (F(8,580) = 56.25, \(p\) \textless{} 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48\%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(8,328) = 5.34, \(p\) \textless{} 0).

@@

Finally, we calculated the ratio of binocular to monocular responses across the three data types from Experiment 1. Figure \ref{fig:BSratios} shows that these ratios are approximately \(\sqrt2\) across the low-to-intermediate contrast range for all three data types. At higher contrasts, we see ratios of 2 or higher for the EEG data, but much weaker ratios near 1 for the pupillometry data. Note that the ratios here are calculated on a per-participant basis and then averaged, rather than being the ratios of the average values shown in Figures \ref{fig:pupildata} and \ref{fig:EEGdata}. A \(3 \times 5\) repeated measures ANOVA on the logarithmic (dB) ratios found a main effect of contrast (F(2.96,121.48) = 6.18, \emph{p} \textless{} 0), no effect of data modality (F(2,82) = 5.55, \emph{p} = 0.01), but a highly significant interaction (F(5.04,206.46) = 6.38, \emph{p} \textless{} 0).

@@

\caption{\label{tab:paramtable}Summary of median parameter values.}
\centering
\begin{tabular}[t]{l|c|c|c|c}
\hline
Data set & Z & k & w & Rmax\\
\hline
Pupillometry & 2.40 & 0.01 & 0.66 & 0.00023\\
\hline
EEG 1F & 2.72 & 0.16 & 0.02 & 0.00261\\
\hline
EEG 2F & 2.23 & 0.06 & 0.01 & 0.00404\\
\hline
Matching & 0.30 & 5.10 & 0.09 & -\\
\hline
\end{tabular}
\end{table}

@@

\hypertarget{declaration-of-interests}{%
\section{Declaration of interests}\label{declaration-of-interests}}

The authors declare no competing interests.
